---
params:
  N:
    label: "Number of Participants"
    value: 200
    min: 0
    step: 1
    input: numeric
  prefix:
    label: "Add a prefix if needed."
    value: ""
    input: text
  name:
    label: "Name of the Room"
    value: "Session_1"
    input: text
title: "Create Rooms for oTree Sessions"
description: |
author:
  - name: Hauke Roggenkamp 
    url: https://www.ibt.unisg.ch/team/hauke-roggenkamp/
    affiliation: Institute of Behavioral Science and Technology
    affiliation_url: https://www.ibt.unisg.ch/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(stringi)
library(magrittr)
library(data.table)
library(knitr)
set.seed(14)
```

This script writes some random tokens that can be used to generate persistent identifiers for  [ROOMS](https://otree.readthedocs.io/en/latest/rooms.html) in oTree. Defining these labels _ex ante_, helps you to match your IDs (the `participant.label`) with the IDs agencies (such as prolific, MTurk or others) provide. This matching procedure is necessary to administer the individual payoffs.

```{r createRoom}
strings <- paste0(params$prefix,
                  stri_rand_strings(n = params$N,
                                    length = 8,
                                    pattern = "[A-Za-z0-9]")) %>%
  data.table()

path <- paste0("../otree/_rooms/", params$name, ".txt")

write.table(strings,
          file = path,
          col.names = FALSE,
          row.names = FALSE,
          quote = FALSE)
```


This is what the first five `participant.label`s you just created look like.

```{r showLables}
setnames(x = strings,
         new = params$name)

kable(head(strings, 5))
```

You stored them in: ``r path`` 

***

_Note that this is a Rmd file you can knit with parameters. This will prompt a UI asking you for the number of participants, whether their labels should contain a custom prefix as well as the name of the room._