---
title: "Pre-processing"
author: Hauke Roggenkamp | 
date: "`r Sys.Date()`"
output:
  github_document:
    toc: false
    toc_depth: 1 
# distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Background

This document belongs to a [project](https://github.com/Howquez/say-it-out-loud) in which participants of an experiment receive an endowment and can decide whether and how much the want to donate to some charity.

Importantly, participants encounter one of two interfaces to communicate their decision: They either type their decision using a classic text input or they make a voice recording that we analyze using a speech-to-text engine. Which type of interface a participant encounters is determined randomly.

You can find the experiment's demo [here](https://ibt-hsg.herokuapp.com/demo).

**Read further if you want to learn how Base64 strings (stemming from recorded voice) is decoded to audio files and then converted to text which is then converted to numeric information.**



# Workflow

1. Run [oTree Experiment](https://ibt-hsg.herokuapp.com/demo) and download the resulting .csv file.
2. Run these lines of code. They will:
    + [decode](#decode) a Base64 string to a .wav file and store it
    + pass these files to a [speech-to-text](#speech-to-text) API (Google)
3. [extract the quantities](#words-to-numbers) from Google's text output
    

# Setup

```{r packages, warning = FALSE}
# install this package once, it'll then install and load all remaining packages
# install.packages("pacman")

pacman::p_load(rmarkdown, knitr, magrittr, data.table, stringr, lubridate, 
               base64enc, googleLanguageR)

# unfortunately, there is one github project we need where the repository's
# name does not match the package's name.
pacman::p_load_gh("fsingletonthorn/words_to_numbers")
library(wordstonumbers)
```

After installing and loading the packages, the data is loaded and stored as a `data.table`.^[According to the authors, `data.table` provides an enhanced version of `data.frame`s that reduce programming and compute time tremendously.]

```{r mostRecentData}
# find most recent file in directory
cFiles <- file.info(list.files(path = "../../data/raw",
                               full.names = TRUE,
                               pattern = ".csv$"),
                    extra_cols = FALSE)
recentCSV <- cFiles[cFiles$mtime == max(cFiles$mtime), ] %>% row.names()
rm("cFiles")
```

Some of the columns contain `base64`-encoded strings that can be decoded to audio files using the `base64enc` package. This implies that we only need these lines of code as well as a single .csv file (in this case ` `r recentCSV` `^[This is the most recent data we have.]) to create and analyze the audio files we are interested in.

```{r readData}
# read data
dt <- data.table(read.csv(file = recentCSV))
```

The data looks approximately^[I changed the order of some selected columns and truncated the long strings for illustrative purposes.] as follows: There are some unique IDs for each participant, a page index, two long strings containing the base64 encoded information and a numeric `share` variable describing contributions made to the charity, as well as many more columns that are not displayed in what follows.

Importantly, there is also a treatment variable called `D_Charity.1.player.voice_interface`. This variable determines whether `D_Charity.1.player.share` or `D_Charity.1.player.voiceBase64` is populated since it describes whether participants face a voice interface or a classic numeric text input.

```{r displayData}
# display relevant columns
dt[, .(participant.id_in_session, participant.code, participant._index_in_pages,
       D_Charity.1.player.voice_interface,
       D_Charity.1.player.comprehensionAudio = D_Charity.1.player.comprehensionAudio %>% str_sub(1500, 1575) %>% paste0("..."),
       D_Charity.1.player.voiceBase64 = D_Charity.1.player.voiceBase64 %>% str_sub(1500, 1575),
       D_Charity.1.player.share)] %>%
        kable()
```

Because the data table is wide (there are `r NCOL(dt)` columns and `r NROW(dt)` rows), variables, that are measured multiple times, are arranged in separate columns. I rearrange it using `melt()`. I then select columns that change over time.

```{r wide2long, warning = FALSE}
# melt, i.e. create long data table
tmp <- melt(data = dt,
            measure = patterns("share$", "voiceBase64$", "comprehensionAudio$"),
            value.name = c("share", "voiceBase64", "baselineBase64"),
            variable.name = "round")

# select ID columns as well as columns that represent variables of interest
long <- tmp[,
            .(session.code,
              participant.code,
              participant.label,
              round,
              voiceInterface = D_Charity.1.player.voice_interface,
              share,
              voiceBase64,
              baselineBase64)]

```

This leaves us with da data.table that has only `r NCOL(long)` columns (and `r NROW(long)` rows) which is more handy to analyze.

Note that I also renamed the variables along the way.^[In a long format, the variables do not correspond to observations in a specific period, which is one of the reasons I can truncate the columns' naming considerably.]


# Decode

The following lines document a for loop that decodes the relevant rows of `voiceBase64` and stores the result in `../../data/wav/`.

```{r convertBase64}
# create file names for the wav files that will be generated using a for loop
long[voiceInterface == 1,
   decisionFilePath := paste0("../../data/wav/",
                       participant.code,
                       "_desicion", # to distinguish between decisions and the baseline
                       round,
                       ".wav")]

long[!is.na(baselineBase64),
   baselineFilePath := paste0("../../data/wav/",
                       participant.code,
                       "_baseline", # to distinguish between decisions and the baseline
                       ".wav")]

# create a vector of "speaking" decisions to loop over
spokenDecisions <- long[!is.na(decisionFilePath) & !is.na(voiceBase64),
                          decisionFilePath]

# run loop to decode and store corresponding files
for(file in spokenDecisions){
        long[decisionFilePath == file,
           base64decode(what = voiceBase64,
                        output = file(decisionFilePath,
                                      "wb"))]
}

# repeat the last two steps for baseline audios
baselines <- long[!is.na(baselineBase64), baselineFilePath]
rm(file)
for(file in baselines){
        long[baselineFilePath == file,
           base64decode(what = baselineBase64,
                        output = file(baselineFilePath,
                                      "wb"))]
}

```


# Speech to Text

To use the voice data, I am using the `googleLanguageR` package and its `gl_speech()` function. This requires you to [authenticate](#authentication). Subsequently, the text is [transcribed](#transcription) which allows us to [detect](#detection) the amount of money participants want to donate.

## Authentication

To use that function, you need to authenticate. Click [here](https://bookdown.org/paul/apis_for_social_scientists/google-speech-to-text-api.html#prerequisites-8) to find out how.

I stored my key in a JSON file and read it via `gl_auth()`. You have to enter the path to your personal key to run the next few lines.

```{r auth}
gl_auth("../../../ibtanalytics-0ba5cc05ef54.json")
```

## Trasincription

### Decision Data

Having done so, I create an empty character vector `spokenDecision` that will be populated using a for loop that calls `gl_speech()`.

```{r gl_speech_1, eval = TRUE}
# initiate empty vector
long[, spokenDecision := as.character("")]

# populate vector in for loop
for(file in spokenDecisions){ # note that there is a difference between spokenDecisions & spokenDecision
        tmp <- gl_speech(audio_source = file, 
                         languageCode = "en", 
                         encoding = "FLAC", 
                         sampleRateHertz = 44100, 
                         customConfig = list('audio_channel_count' = 1))
        
        transcript <- tmp$transcript[,1]
        
        long[decisionFilePath == file,
           spokenDecision := transcript]
}

# display vector
kable(long[!is.na(voiceBase64),
           .(spokenDecision)])
```

The quality of the transcription is rather bad, but good enough as each of the numbers contained is correct. To listen to the corresponding audio files, you have to browse to `../../data/wav/`. The original text of the first three entries were _I transfer [1, 3, 6] point(s)_. The last two entries should read _I donate [7, 9] points_.

### Baseline Data

Because I also have the baseline recordings where participants say _I have read and understand the instructions_, we can transcribe them as well. Since I do not intend to analyze the content, I set the corresponding code chunk option such that the following few lines are not evaluated. This implies, that I will not use the information in what follows until some voice analytics is applied.

```{r gl_speech_2, eval = TRUE}
# initiate empty vector
long[, spokenBaseline := as.character("")]

# populate vector in for loop
for(file in baselines){
        tmp <- gl_speech(audio_source = file, 
                         languageCode = "en", 
                         encoding = "FLAC", 
                         sampleRateHertz = 44100, 
                         customConfig = list('audio_channel_count' = 1))
        
        transcript <- tmp$transcript[,1]
        
        long[baselineFilePath == file,
           spokenBaseline := transcript]
}

# display vector
kable(long[!is.na(baselineBase64),
           .(spokenBaseline)])
```


# Words to Numbers

To extract the quantities described in the  `spokenDecision` vector, one can use the `wordstonumbers` package and apply the corresponding function (`words_to_numbers`) to it. As this just replaces the spelled number with a digit, one also has to remove all the characters from the respective strings. I do so, and write the result as a numeric into a variable called `spokenDecision2text`. 

**Note** that for each of these transcriptions, Google bills 15s.^[Also note that you have a free monthly contigent of 60 minutes.]

```{r speech2text, warning = FALSE}
speechtotext <- apply(X = long[,.(spokenDecision)],
                      MARGIN = 1,
                      FUN = words_to_numbers) %>%
        str_replace_all(pattern = "\\D",
                        replacement = "") %>%
        as.numeric()

long[, spokenDecision2text := speechtotext]
```

Ultimately, this variable can be blended with `share`, such that we have one column that contains the donations of all participants. This variable, I call this variable `donation`, is our primary outcome. We expect it to differ between the two treatments.^[To put it differently, `spokenDecision2text` and `share` should differ.]

```{r createOutcome}
# create donation variable
long[, donation := share]
long[!is.na(spokenDecision2text), donation := spokenDecision2text]
```
# Result

Finally, one can select some technical covariates (`techC`) as well as covariates stemming from the questionnaire (`questC`) that do not change within the session and merge them to the `long` data.table.

```{r technicalCovariates}
# create DT with technical covariates from first period
# (assuming they are constant across rounds)
techC <- dt[,
            .(session.code,
              participant.code,
              participant.label,
              longitude        = D_Charity.1.player.longitude,
              latitude         = D_Charity.1.player.latitude,
              ipAddress        = D_Charity.1.player.ipAddress,
              screenWidth      = D_Charity.1.player.width,
              screenHeight     = D_Charity.1.player.height,
              devicePixelRatio = D_Charity.1.player.devicePixelRatio,
              userAgent        = D_Charity.1.player.userAgent)]

# merge
long <- data.table::merge.data.table(x = long,
                                     y = techC,
                                     by = c("session.code",
                                            "participant.code",
                                            "participant.label"))
```

```{r questionnaireCovariates}
# create DT with covariates from questionnaire
questC <- dt[,
             .(session.code,
               participant.code,
               participant.label,
               Q1 = D_Charity.1.player.comp_1)]

# merge
long <- data.table::merge.data.table(x = long,
                                     y = questC,
                                     by = c("session.code",
                                            "participant.code",
                                            "participant.label"))
```

The resulting data looks as follows.

```{r displayFinalData}
display <- long
display[!is.na(voiceBase64), voiceBase64 := voiceBase64 %>% str_sub(1500, 1575) %>% paste0("...")]
display[!is.na(baselineBase64), baselineBase64 := baselineBase64 %>% str_sub(1500, 1575) %>% paste0("...")]
kable(display)
```

